How to be a good member of a scientific software community [Article v\versionnumber]

### 2022-01-17 response-letter.tex

\section{Reviewer A}

1) My biggest concern is simply that the people who ought to read it,
won't do so (even if directed) because it is longer than what users
typical want to read when they are in search of a fix to their
problem. In this the article shares the same shortcoming with the
well-known online essays [1] and [2].

I'd like a condensed summary of about one paragraph (maybe as a
box) that I can just paste in an email or put on a website,
together with a link to the full version.

\response{I have added references to the Raymond and Tatham essays -- I'm
somewhat embarrassed I didn't think to add them myself, since they clearly
influenced my thinking. I would argue that the checklists make
a good short punchy summary, as the reviewer suggested. However, I agree having
a plain-text version wouldn't hurt, so I've added a ``Summary'' paragraph before
the checklists.}

@@

On the other hand, if code is advertised (presumably with the
expectation of citations or some other benefit to the developer)
then I agree with the author that developers owe their users some
of their time.

\response{My focus was mostly on code that is intended to be used. I've added a
paragraph at the end of the section for developers laying out my thoughts on
this distinction.}

3) "Please open a PR" (or merge request) is a phrase that can strike
fear in the heart of a novice. It might be helpful to demystify
this important part of collaborative software development. Maybe
the a PLoS Comp Biol article [5] or something similar that touches
upon basic software engineering that could provide more guidance.

\response{I agree, and I'd argue that the relevance here is that it falls to the
developers to provide a less intimidating way to contribute. I'd rather not talk
about specific software engineering technologies here, though, just because I
feel like it's a topic I couldn't do justice to without doubling the length of
the paper.}

### 2022-01-17 paper.tex
added ref to Lindahl's paper

This article is written from a the perspective of a developer of academic
open-source scientific software, based on my personal experience as a developer
and user. I released my first piece of open source code, wham \cite{WHAM}, a
tool for analyzing umbrella sampling simulation,  during the summer of 2000.  My
group released LOOS\cite{Grossfield-2009, LOOS-JCC}, a suite of tools for
developing new tools to analyze molecular dynamics simulations, in 2008. Over
the years, I've interacted with hundreds of people who've tried to use one
package or the other. Most of these interactions have been positive, some
overwhelmingly so. Others were not. This article is an attempt to distill my
experiences into a concise set of recommendations.

Much of what I suggest is universal, but some points --- like the
reminder that the people providing the support are likely volunteers --- are
not.  Regardless, we view the interaction between developers and users as an
informal social contract, where each has obligations and expectations for the
other. Obviously, every software community is unique, and many have specific
conventions for interaction, but we hope that the recommendations we make are at
the very least a good starting point for new users of scientific software.

@@

Doing these things is valuable from a practical perspective, as discussed
elsewhere \cite{Heister-2013, Laayouni-2016}; making it easy for people to
interact with us efficiently, even if it costs us some effort up front, will
make those interactions more pleasant. Even setting that aside, it's the right
way to do science --- FAIR principles for software development
\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
That means documenting it well and supporting people who try to use it.

>>>>

This article is written from a the perspective of a developer of academic
open source scientific software, based on my personal experience as a developer
and user. I released my first piece of open source code, wham \cite{WHAM}, a
tool for analyzing umbrella sampling simulation,  during the summer of 2000.  My
group released LOOS\cite{Grossfield-2009, LOOS-JCC}, a suite of tools for
developing new tools to analyze molecular dynamics simulations, in 2008. Over
the years, I've interacted with hundreds of people who've tried to use one
package or the other. Most of these interactions have been positive, some
overwhelmingly so. Others were not. This article is an attempt to distill my
experiences into a concise set of recommendations.

Much of what I suggest is universal, but some points --- like the
reminder that the people providing the support are likely volunteers --- are
not.  Regardless, I view the interaction between developers and users as an
informal social contract, where each has obligations and expectations for the
other. Obviously, every software community is unique, and many have specific
conventions for interaction, but we hope that the recommendations we make are at
the very least a good starting point for new users of scientific software.

@@

Doing these things is valuable from a practical perspective, as discussed
elsewhere \cite{Heister-2013, Laayouni-2016, Wallqvist-2019}; making it easy for
people to interact with us efficiently, even if it costs us some effort up
front, will make those interactions more pleasant. Even setting that aside, it's
the right way to do science --- FAIR principles for software development
\cite{Lamprecht-2020} prioritize not just releasing code, but making it useable.
That means documenting it well and supporting people who try to use it.
